WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis

### 2025-07-06 docs/manuscript_GNM.qmd
Results comments addressed (initial stage).. continue results and discussion

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd,patchwork)
source(here::here("functions/extractor_metareg.R"))
```

@@

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details the study's objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022. The keyword search query was designed as a three-block query to included terms related to the predictor/treatment (("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), our population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and to limit it to only experimental studies only ("experiment\*" OR "manipulat\*"). The complete search strings tailored to each specific platform are provided in @figsupp-string. We searched through titles, abstracts and keywords in English-language only with no restrictions on publication date or study type. We validated our search using an initial library of `{r} nrow(own_library)` relevant articles that we had previously identified (@figsupp-own-library), ensuring all were retrieved. Additionally, we used the R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019).

@@

All titles and abstracts were equally allocated and screened independently by atleast two screeners (SD, TR, JMGS) using the R package `revtools` following a pre-designed decision tree (@figsupp-decision-tree). We included all studies on birds that experimentally manipulated the green nest material (both aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed the title-and-abstract screening and were further evaluated during the full-text screening. Each article was screened by a single screener (SD, TR, JMGS), with another screener verifying the procedure. We used the same criteria as above (see @figsupp-decision-tree) with an additional evaluation of whether the variable of interest in the article was in our list of fitness proxies of interest (see @figsupp-fitness-proxies). `{r} nrow(excluded_fulltext)` articles were excluded. All conflicts in screening decisions (\~1% disagreement for title-and-abstract screening, and none for full-text screening) were resolved through discussion.

To update our systematic lietrature search, we repeated our search on 12.08.2024. This updated search found `{r} nrow(unique_reference_list_repeat)` unique articles after removing ``` {r} nrow(full_reference_before_deduplication_repeat)-``nrow(unique_reference_list_repeat) ``` duplicates. Only one observer (SD) performed the title-and-abstract and full-text screening using the same decision trees as before and found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` additional eligible article. Lastly, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` additional eligible articles. In all, X <!--# Fill the no. of articles that are included --> articles were eligible for our meta-analysis (more details in @figsupp-prisma-flowchart).

@@

excluded_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="exclude")
SAonly_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="senstivity-analysis" | proxy_decision=="sensitivity-analysis")
```

The data of each article was extracted evenly by one of the three authors (SD, TR, JMGS), with another screener verifying the procedure. We extracted the primary data (i.e., measures of central tendency, variance and sample size) from text whenever possible. In cases where primary data was only available in figures (e.g., boxplots), we used `metaDigitise` v.1.0.1 (@pick2018). If primary data was not available, we extracted any inferential statistics (t value, F value) that could be transformed into an effect size. In cases where the values could not be extracted from the article but datasets were provided, we calculated the means, SD and sample sizes from the provided raw data. When data was presented in different sources (e.g. in a boxplot and as an inferential statistics), we prioritized the source having the maximum resolution (the boxplot, in this example).

To obtain missing data (e.g., non-reported results, missing sample sizes, etc.) and to locate grey literature (i.e. unpublished data), we wrote to the corresponding author of the articles that passed full-text screening using a standardized email template (\@figsupp-email). We contacted four authors for missing data, while all corresponding authors were asked for any potential unpublished data that could be included in our meta-analysis. All four authors responded to our email but only two could provide the data as requested; one did not have the data any more and one did not give a conclusive response. In addition, one author provided data for an unpublished study (see description in @figsupp-unpublished-study) and one of the authors of this study (MO) provided data for an unpublished study. In all, we extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

We extracted the means, variances and sample size for each control-treatment pairwise comparison. Controls included non-aromatic materials such as grass or moss, lesser amounts of aromatic material, or no material at all. Treatments consisted of aromatic materials added by authors, birds, or both. In cases comparing non-aromatic material to a blank nest, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator.

For each study we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, fitness proxy measured) and information about the extracted data (location or source of the data recorded). We categorized the fitness proxy measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load and phenology. For parasites or pathogens, we categorised them into arthropods or micro-organisms. For a complete list of variables extracted see @figsupp-variables . We did not include some traits in all our analyses becaause it was unclear how they related to fitness (k=`{r} nrow(excluded_proxies)` , n=`{r} length(unique(excluded_proxies$paper_ID))`;see @figsupp-excluded-proxies for the list and reason of exclusion). We also had to exclude some effect sizes (n=?, k=?) because of the low sample size per group (less than x).

@@

To estimate the effect of green nest material addition on fitness across birds, we used two effect sizes: lnRR and SMDH. Thus, throughout, we fitted two separate sets of model for each effect size. First, we fitted multilevel intercept-only meta-analytical model to test the overall effect, its uncertainty and heterogeneity. We used the `rma.mv` function from the `metafor` package using the 'REML' method and a *t* distribution for calculating test statistics and confidence intervals for the fixed effects. We also conducted a sensitivity analysis including some traits for which the relationship with fitness was unclear, to test the robustness of our results (@figsupp-MLMA-excluded; k=`{r} nrow(SAonly_proxies)` , n=`{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly_proxies). For all intercept-only models, we estimated the total amount of absolute heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) following the pluralistic approach suggested by @yang2025.

<!--# You have to correct the numbers for contingency table. It does not include the one data point from unpublished results -->

@@

To test the relative importance of the two main hypotheses suggested for green nest material: Courtship hypothesis and Parental care hypothesis (combining nest protection and drug hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. For this and all meta-regressions, we estimated $R_{marginal}^2$, which corresponds to the percentage of heterogeneity explained by the moderators (@nakagawa2012).

@@

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of green nest material on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (n=?, k=?). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with with the entire dataset set using bird species as a moderator (levels: Cyanistes caeruleus, Sturnus unicolor, Tachycineta bicolor, Sturnus vulgaris, Parus major, Passer cinnamomeus, Buteo buteo) to understand if the effect of green nest material on fitness depended on the system being studied. To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase) we ran a meta-regression with Time of addition of green nest material as the moderator. We explored the effect of green nest material on fitness depending on the type of trait studies (levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with the type of experimental design as a moderator.

@@

Publication bias occurs when a subset of research findings, most often statistically non-significant ones, are less likely to be published or being published as compared to the rest. We tested for evidence of publication bias, both small-study and decline effects following @nakagawa2021. Publication bias tests were performed for both lnRR and SMDH separately, and using the same random effect structure as reported above. To test for small-study effects, we ran a multilevel meta-regression, using the inverse of the effective sample size as a moderator. To test for decline effects (also known as time-lag bias; @sanchez-tojar2018), we ran multilevel meta-regression year of publication (mean-centered) as a moderator.

@@

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` geographical locations. About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes came from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`). Our dataset included two unpublished studies (k=XX effect sizes), with the remaining studies being published between `{r} min((dataset_analysis$year_publication),na.rm = T)` and `{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} as.numeric(format(mean(ES_per_study$n), digit=3))` ± `{r} as.numeric(format(sd(ES_per_study$n), digit=2))`(range: `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)`) and the median and mean sample size per effect size were `{r} median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))` and `{r} as.numeric(format(mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)), digit=3))` respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`--`{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%select(paper_ID,effective_n_control,effective_n_experiment, lnRR, SMDH)%>%filter(effective_n_experiment<2|effective_n_control<2))` effect sizes had a sample size of less than 2 in each group).

@@

```{r overall-results}
intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))
bivariate_model<-readRDS(here::here("sensitivity_analysis/bivariate_model.rds"))
```

Both lnRR and SMDH intercept-only multilevel meta-analyses showed an increase in fitness in the nests with an increase in green nest material compared to control nests; however, the overall effect was only statistically significant for SMDH.

>>>>

# Abstract

According to PRISMA Checklist.. I must:

-   Summarise the aims and scope of the review

-   Describe the data set

-   State the results of the primary outcome

-   State conclusions

-   State limitations

# Introduction

Nest and nest-building are fundamental across a diverse range of animals, including insects (e.g., termites, spiders), mammals (e.g., platypus, rodents), and perhaps more famously, birds. Nests provide structural support for eggs, parents, and offspring, and protect them against predators and environmental pressures, and are thus key to reproductive success and survival [@mainwaring2014; @deeming2023; @hansell2000]. While most birds use organic plant materials such as twigs, branches, grass, and moss, some incorporate inorganic items, including stones, feathers, snake skins and even anthropogenic materials like plastic and cigarette butts [@suárez-rodríguez2013; @sergio2011; @sheard2024]. These non-structural additions suggest that nests are more than just physical shelters; they can be complex tools shaped by both natural and sexual selection for multiple functions. Indeed, nests can play a key role in sexual display, and features such as size have been shown to act as sexually selected signals of parental quality in many bird species [@moreno1994; @hoi1996], even across species [@soler1995]. Nest-building behaviour is increasingly recognized as an expression of the extended phenotype (reviewed by @moreno2012), and phylogeny has been shown to be an important factor explaining both mean and within-species variation in nest morphology [@perez2023]. Yet, despite hundreds of scientific articles studying nests and nesting behaviour in birds being published each year, much remains unknown.

An intriguing nesting behaviour observed in some bird species is the addition of fresh, often aromatic, plant material to the nest. This green nest material (GNM) ---leaves, sprigs, herbs and branches--- is typically added from the start of nest building until egg laying, although some species keep adding it even after the nest is finished. Importantly, contrary to the structural nest materials (e.g., twigs, branches), GNM is thought to be selected from a small, non-random subset of local flora [@clark1985; @garrido-bautista2023], and often characterized by high levels of aromatic volatile compounds (commonly used by songbirds: @clark1985) or secondary metabolites such as oils and resins (commonly used by raptors: @ontiveros2007). Collecting and constantly replenishing GNM requires extra foraging flights, sometimes even outside the territory [@mennerat2009]. The costs associated with nest building [@mainwaring2013], and these additional efforts suggest that GNM is unlikely to be gathered opportunistically and suggests a possible adaptive significance.

Several hypotheses have been proposed to explain the adaptive function of GNM [@dubiec2013; @scott-baumann2015]. First, the courtship hypothesis suggests that mostly males use GNM as visual or olfactory signals to attract females or advertise nest site occupancy, particularly in cavity-nesting species since visual cues inside the nest are limited. This hypothesis traces back to the early natural history observations by @kessel1957, who noted that male European Starlings (*Sturnus vulgaris*) carried fresh plant material to their nests while singing and displaying to females. Subsequent studies in both European and Spotless Starlings (*Sturnus unicolor*) supported this view that GNM may serve as a sexually selected signal of male quality by showing that (i) unpaired males conspicuously deliver GNM during courtship, stopping right after the onset of egg laying [@brouwer2004; @veiga2006], and seemingly increasing this behaviour in the presence of potential mates [@eens1993], and (ii) females often prefer nests with fresh GNM. The courtship hypothesis was later extended to explain GNM addition solely by females in species like Blue Tits (*Cyanistes caeruleus*), therefore suggesting that GNM might serve as a sexually selected signal of female quality [@tomás2013].

Second, and the most widely tested one, is the nest protection hypothesis, which proposes that the volatile compounds in GNM reduce the amount of parasites and pathogens in the nests, enhancing offspring survival. This hypothesis traces back to a comparative study in raptors (Falconiformes) that proposed it and provided indirect evidence for it by comparing the nest reuse, diet composition and GNM presence [@wimberger1984]. Soon after, a preference for GNM containing insect-repellent and antimicrobial volatiles was experimentally confirmed in European Starlings, therefore confirming two of the three key criteria for evaluating the nest protection hypothesis (i.e., selective plant use, bioactive chemical properties and demonstrable adaptive benefits; @clark1988). Observational and experimental studies on raptors [@ontiveros2007; @dykstra2009] and passerines [@dawson2004; @lafuma2001; @mennerat2009b; @quiroga2012] have shown mixed results, and in some cases, GNM has been shown to improve offspring condition without reducing the amount of parasites in the nests [@gwinner2000], which led to the development of the drug hypothesis. The drug hypothesis suggests that the volatiles present in GNM directly enhance offspring immune function and/or health independent of parasite or pathogen control [@gwinner2000; @mennerat2009a]. Thus, while both hypotheses suggest an adaptive benefit of GNM by enhancing offspring survival, the drug hypothesis emphasizes a pharmacological mechanism, whereas the nest protection hypothesis emphasizes parasite and pathogen repellency. Importantly, these two hypotheses are not mutually exclusive unless the improvement in offspring health can be proved to be unrelated to a reduction in parasite or pathogen load.

Last, several additional adaptive hypotheses or functions of GNM that do not refer to its aromatic properties have been proposed. These include benefits such as enhancing the crypticity of the nest [@skutch1957], providing insulation or sun protection for eggs and offspring [@mertens1977], or simply fulfilling a structural function [@lyons1987]. However, these alternative functional roles have received little support or attention. In all, despite decades of research, abundant observational and experimental studies, and two crucial narrative reviews [@dubiec2013; @scott-baumann2015], we still do not know the overall adaptive significance of GNM, its generality across bird species, and the relative importance of each of the suggested mechanisms. Given that nest building is widespread across taxa ---including but not limited to birds--- and is increasingly recognized as an expression of the extended phenotype, a quantitative synthesis of this field is both timely and potentially far-reaching in its implications.

To address this gap, we conducted a pre-registered systematic review and meta-analysis of experiments testing the effect of GNM on various fitness proxies across birds. Our meta-analysis explicitly evaluates the overall evidence for the adaptive significance of GNM, as well as assesses the relative importance of the main suggested hypotheses: the courtship hypothesis and parental care hypothesis (i.e. combining both the nest protection and drug hypotheses due to their non-exclusivity). In addition, we explore the observed context-dependency (i.e., high heterogeneity among effect sizes) via meta-regressions. Specifically, we tested whether the effect of GNM varied depending on the type of parasite (arthropods vs. microorganisms), the timing of GNM addition (before hatching, after hatching, or continuously), the experimental design (aromatic vs. non-aromatic vs. no added material), the focal species, and the type of fitness proxy investigated (e.g., physiology, morphology, reproduction). Our meta-analysis of experimental studies is the first quantitative synthesis testing the causal relationship between GNM and fitness across birds.

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd,patchwork,english)
source(here::here("functions/extractor_metareg.R"))
```

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details our objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist and @figsupp-prisma-flow). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

@@

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/01_systematic_search/02_reference_data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/01_systematic_search/02_reference_data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/01_systematic_search/03_title_and_abstract_screening/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/01_systematic_search/04_search_repeat_12082024/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/01_systematic_search/04_search_repeat_12082024/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/01_systematic_search/04_search_repeat_12082024/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/01_systematic_search/03_title_and_abstract_screening/excluded_fulltext.csv")
```

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022 (see update below). The keyword search query was designed as a three-block query to include terms related to our treatment of interest (("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and to limit the search results to experimental studies only ("experiment\*" OR "manipulat\*"). The complete search strings tailored to each specific platform are provided in @figsupp-string. We searched through titles, abstracts and keywords in English-language only with no restrictions on publication date or study type. We validated our search using an initial library of `{r} nrow(own_library)` relevant articles that we had previously identified (@figsupp-own-library), ensuring all were retrieved. Additionally, we used the R package 'litsearchr' v1.0.0 [@grames2019], which did not yield any new relevant terms, to confirm our keyword selection. After removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 [@westgate2019], we found `{r} nrow(unique_reference_list)` unique records.

@@

All titles and abstracts were equally allocated and screened independently by atleast two screeners (SD, TR, JMGS) using the R package `revtools` following a pre-designed decision tree with all inclusion/exclusion criteria (@figsupp-decision-tree). We included all studies on birds that experimentally manipulated the GNM (both aromatics or herbs). `{r} tools::toTitleCase(as.character(english::english(nrow(selected_abstract_screening))))` articles passed the title-and-abstract screening and were further evaluated during the full-text screening, where each article was independently screened by a single screener (SD, TR, JMGS), with another screener verifying the procedure. We used the same criteria as above (@figsupp-decision-tree) with an additional evaluation of whether the variable of interest in the article was in our list of fitness proxies of interest (see @figsupp-fitness-proxies). `{r} paste0(toupper(substring(x <- as.character(english::english(nrow(selected_abstract_screening) - nrow(excluded_fulltext))), 1, 1)), substring(x, 2))` passed the full-text screening and were, thus, eligible for our meta-analysis. All conflicts in screening decisions (\~1% disagreement for title-and-abstract screening, and none for full-text screening) were resolved through discussion.

We updated our systematic literature search by repeating our search on 12.08.2024. We found `{r} nrow(unique_reference_list_repeat)` unique articles after removing `{r} nrow(full_reference_before_deduplication_repeat)-nrow(unique_reference_list_repeat)` duplicates. Only one observer (SD) performed the title-and-abstract and full-text screening using the same decision trees as before (@figsupp-decision-tree), and found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` additional eligible article for our meta-analysis. Lastly, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` additional eligible articles through other sources (i.e., identified through a poster by a collaborator and through an independent recommendation by a colleague). In all, `{r} nrow(selected_abstract_screening)-nrow(excluded_fulltext)+nrow(selected_abstract_screening_repeat)` articles were eligible for our meta-analysis (more details in @figsupp-prisma-flowchart).

@@

excluded_proxies<- merged_data_extraction %>%
  filter(fulltext_screening == "included" |
         is.na(fulltext_screening) |
         fulltext_screening == "NA") %>%
  filter(proxy_decision== "exclude")

dataset_analysis <- read_csv("data/04_data_analysis/dataset_analysis.csv", show_col_types = F)

SAonly_proxies<- dataset_analysis %>% filter(proxy_decision=="sensitivity-analysis")

inferential_only<- dataset_analysis%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_analysis%>%filter(proxy_decision=="contingency table")

raw_data<-c("GNM_001_rep_table3","GNM_010","GNM_069","GNM_237") 
#GNM_018 also provides open data but we ended up not needing it

published_studies <- dataset_analysis %>%
  filter(!str_detect(paper_ID, "Unpublished"))

unpublished_studies <- dataset_analysis %>%
  filter(str_detect(paper_ID, "Unpublished"))

The data of each article was extracted evenly by one of the three authors (SD, TR, JMGS), with another screener verifying the procedure. We prioritised extracting the primary data (i.e., measures of central tendency, variance and sample size) from text and tables, or, if that was not possible, from figures (e.g., boxplots) using the R package `metaDigitise` v.1.0.1 (@pick2018). If primary data were not available, we extracted inferential statistics such as *t* and *F* values (k = `{r} nrow(inferential_only)`, n = `{r} length(unique(inferential_only$paper_ID))`) or descriptive statistics such as contingency tables (k = `{r} nrow(contingency_table)`, n = `{r} length(unique(contingency_table$paper_ID))`), that could be transformed into an effect size. Last, if the previous steps failed but the article has openly shared datasets (n = `{r} length(raw_data)`), we calculated the means, SD and sample sizes from them. **When data was presented in different sources (e.g. in a boxplot and as an inferential statistics), we prioritized the source having the maximum resolution (the boxplot, in this example).**

We extracted the means, variances and sample size for each control-treatment pairwise comparison. Controls included non-aromatic materials such as grass or moss, lesser amounts of aromatic material, or no material at all. Treatments consisted of aromatic materials added by authors, birds, or both. For cases that compared non-aromatic material to where authors did not add any additional nest material and any material added by the birds was removed, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator. For each study, we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, fitness proxy measured) and the source from which we extracted data (e.g., figures, tables, raw data files, or supplementary materials). We categorised the fitness proxies measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load, and phenology. For studies testing whether GNM affects the number of parasites or pathogens found in offspring and/or nests, we categorised them into arthropods or microorganisms. For a complete list of variables extracted see @figsupp-variables . We did not include some traits in all our analyses because it was unclear how they related to fitness (k = `{r} nrow(excluded_proxies)` , n = `{r} length(unique(excluded_proxies$paper_ID))`; see @figsupp-excluded-proxies for the list and reasons of exclusion).

To obtain missing data (e.g., non-reported results, missing sample sizes, etc.), we contacted four authors. All four authors responded to our email, but only two provided the data as requested; one did not have the data any more, and one did not give a conclusive response. Additionally, to locate grey literature (i.e., unpublished data), we wrote to the corresponding author(s) of all the articles that passed full-text screening (n = `{r} (nrow(selected_abstract_screening) - nrow(excluded_fulltext))`) to ask for any potential unpublished data that could be included in our meta-analysis. One corresponding author provided data for an unpublished study (see description in @figsupp-unpublished-study) and one of the authors of this study (MO), provided data for another unpublished study. In all, we extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies. Author correspondance was performed using a standardised email template (@figsupp-email). In all, we extracted data for `{r} length(unique(published_studies$paper_ID))` published articles (k = `{r} nrow(published_studies)`) and `{r} length(unique(unpublished_studies$paper_ID))` unpublished studies (k = `{r} nrow(unpublished_studies)`).

@@

To estimate the effect of GNM addition on fitness across birds, we used two effect sizes: lnRR and SMD(H). Thus, throughout, we fitted two separate sets of model for each effect size. First, we fitted multilevel intercept-only meta-analytical model to test the overall effect, its uncertainty and heterogeneity. We used the `rma.mv` function from the `metafor` package using the 'REML' method and a *t* distribution for calculating test statistics and confidence intervals for the fixed effects. We also conducted a sensitivity analysis including some traits for which the relationship with fitness is unclear or debated in the literature (e.g. sex ratio, testosterone levels, etc.), to test the robustness of our results (@figsupp-MLMA-excluded; k = `{r} nrow(SAonly_proxies)` , n = `{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly-proxies). For all intercept-only models, we estimated the total amount of absolute heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) following the pluralistic approach suggested by @yang2025.

@@

To test the relative importance of the two main hypotheses suggested for GNM: Courtship hypothesis and Parental care hypothesis (combining nest protection and drug hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. For this and all meta-regressions, we estimated $R_{marginal}^2$, which corresponds to the percentage of heterogeneity explained by the moderators [@nakagawa2012].

@@

```{r}
parasite_pathogen<-dataset_after_cleaning%>%filter(trait_type=="Parasitic_and_pathogenic")
```

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of GNM on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (k = `{r} nrow(parasite_pathogen)`, n = `{r} length(unique(parasite_pathogen$paper_ID))`). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with the entire dataset set using bird species as a moderator (levels: *Buteo buteo*, *Cyanistes caeruleus*, *Passer cinnamomeus*, *Parus major*, *Sturnus unicolor*, *Sturnus vulgaris*, *Tachycineta bicolor*) to understand if the effect of GNM on fitness depended on the system being studied. To test the effect of time of addition of GNM on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase), we ran a meta-regression with Time of addition of GNM as the moderator. We explored the effect of GNM on fitness depending on the type of trait studies (levels: physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between GNM and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with the type of experimental design as a moderator.

@@

Publication bias occurs when a subset of research findings, most often statistically non-significant ones, are less likely to be published or being published as compared to the rest. We tested for evidence of publication bias, both small-study and decline effects following @nakagawa2021. Publication bias tests were performed for both lnRR and SMD(H) separately, and using the same random effect structure as reported above. To test for small-study effects, we ran a multilevel meta-regression, using the inverse of the effective sample size as a moderator. To test for decline effects (also known as time-lag bias; @sanchez-tojar2018), we ran multilevel meta-regression year of publication (mean-centered) as a moderator.

@@

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` geographical locations (@fig-distribution). About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes came from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`). Our dataset included `{r} length(unique(unpublished_studies$paper_ID))` unpublished studies (k = `{r} nrow(unpublished_studies)`), with the remaining studies being published between `{r} min((dataset_analysis$year_publication),na.rm = T)` and `{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} sprintf("%.1f", mean(ES_per_study$n))` ± `{r} sprintf("%.1f",sd(ES_per_study$n))` (range: `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)`) and the median and mean effective sample size (i.e. after accounting for shared-group non-independence) were `{r} sprintf("%.1f", median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)))` and `{r} sprintf("%.1f",mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)))`, respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`--`{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%filter(n_experiment<5|n_control<5))` data points had a sample size of less than 5 nests in each group).

![**Geographical distribution of the 16 study population in this meta-analysis:** Point size reflects the number of effect sizes coming from the location and colour denotes species (see legend), with an inset zooming on the cluster of distribution.](figures/Figure_Distribution.png){#fig-distribution}

@@

```{r overall-results}
intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))
bivariate_model<-readRDS(here::here("model/sensitivity_analysis/bivariate_model.rds"))
```

Both lnRR and SMD(H) intercept-only multilevel meta-analyses showed an average increase in fitness in the nests with an experimental increase in GNM compared to control nests; however, the overall effect was only statistically significant for SMD(H).
